%% ============================================================
\section{Conclusion}\label{sec:conclusion}
%% ============================================================

Loop invariant generation has been studied for four decades almost exclusively as a target-directed task, yet the verification target it presupposes is scarce in practice and is, for loops, typically the exit-time shadow of the very invariant being sought.
We reframed the problem as \emph{closed-book inductive loop summary generation}---characterize the loop, not the goal---and showed that the missing ingredient for learning this task is not a verifier bit but a dense, execution-grounded, audited reward.
\sam{} supplies it with a deliberately small contract: three synthetic negative-candidate families, conservative construction guards, Houdini-pruned survival, trace-unit accounting, and $r=0.5\,\mathrm{base}+0.5\,\mathrm{marginal}$ with no auxiliary scoring patches.  The 10-program deterministic discrimination harness checks known-quality rankings, and the measured full-suite audit found no hard collision among 486,302 negative witnesses from 366 programs.
The pipeline is deliberately modular---sampler, a reward image with Frama-C/WP bundled, and independent inference that performs no execution sampling and can run verdict-driven multi-round refinement---so that any RL trainer can optimize against it and any deductive toolchain can consume its output.

Beyond the immediate system, we see the two central ideas as portable.
Audited synthetic negatives generalize to domains where construction-specific evidence is cheaper than proving optimality; keeping assumptions and collision checks explicit is preferable to compensating for bad labels in the scoring formula.
Future work includes scaling summaries to multi-loop and heap-manipulating programs~\citep{liu2024towards}, richer assertion languages (quantifiers, arrays), and closing the loop with the specification-generation literature: a tight loop summary is, after all, the specification.
