%% ============================================================
\section{Experiments}\label{sec:experiments}
%% ============================================================

\subsection{Setup}
\label{sec:exp:setup}

\paragraph{Benchmarks.}
We evaluate on 366 C loop programs in two suites.
\textbf{Linear} (316 programs) contains single-loop affine programs in the style of SV-COMP loop tracks~\citep{svcomp}; roughly 40\% use nondeterministic guards or branches (\texttt{unknown()}), exercising the sampler's soundness rescues.
\textbf{NLA} (50 programs) contains the standard non-linear arithmetic suite~\citep{nguyen2012using,nla2019}: polynomial identities (cubes, geometric sums), gcd/extended-gcd, and division-by-subtraction algorithms whose summaries require products and higher-degree terms (e.g.\ $y = 3n^2 + 3n + 1$, $ps - rq = 1$, $z = xy$).
Every program carries a trailing assertion; it is used \emph{only} for final end-to-end checking and is hidden from the sampler and the generator throughout.

\paragraph{Backends and configuration.}
Positives/negatives are sampled with 12 runs per seed under the canonical 20-tier sweep; inductiveness and final certification use Frama-C/WP with Z3 (30\,s per goal, Typed memory model); reward weights and caps are the defaults of $\S$\ref{sec:method:reward}.
The reward runs as a slim container (no prover, no GPU dependencies) and scores a rollout group in the low seconds per program, dominated by Houdini when the prover is enabled.

\subsection{Adversarial Validation of the Reward}
\label{sec:exp:gates}

A reward for RL is only as good as its worst exploit, so before any policy training we validate the reward itself with three adversarial regression gates run across both suites.

\paragraph{Gate 1: discrimination.}
For each program we score hand-constructed families: \emph{gold} (true tight summaries), \emph{loose} (sound but weak, e.g.\ guard restatements, slack boxes), and \emph{adversarial} (mega-conjunctions, split sprays, gold-plus-spray, sample-extreme boxes, pointwise farms).
The gate requires the strict quality order $\mathrm{gold} \ge \mathrm{loose} > \mathrm{hack}$ pointwise, memorizers at least $0.15$ below gold, and sample-extreme boxes collapsing on the holdout seed.
\textbf{Result: 0 violations.}

\paragraph{Gate 2: gold-free hack sweep.}
Independent of any gold reference, parametric adversary families sweep their capacity (clause counts, atom counts, farm sizes) against every program; the gate requires that no adversary configuration attains the reward of a true summary on any program.
\textbf{Result: 0 violations.}

\paragraph{Gate 3: mislabel audit.}
Truthfulness of negatives is audited directly: every emitted negative trace unit is re-checked against exhaustive forward reachability within the audited window, and any reachable state labeled negative is a hard failure (it would penalize true invariants).
\textbf{Result: 0 hard mislabels} across all 366 programs, including the nondeterministic subset where negatives come only from the rescue constructions.

Together the gates certify the two properties a hack-resistant verifiable reward needs: \emph{soundness} (no true invariant is punished) and \emph{discrimination} (no known exploit outscores truth).
They run as a regression suite, so any future change to the sampler or reward must re-pass them.

\subsection{Policy Training and End-to-End Generation}
\label{sec:exp:rl}

%% TODO(experiments): fill in once RL runs complete.
%% - Base models: Qwen-2.5-Coder-7B (local vLLM), GPT-5-nano (API reference)
%% - Trainer: GRPO against the HTTP reward service; n rollouts/group = 8
%% - Metrics: (a) mean final reward (summary tightness) on held-out programs;
%%   (b) Frama-C-certified summary rate (closed-book);
%%   (c) downstream: fraction of benchmark assertions discharged by the
%%       generated summary alone (I ∧ ¬B ⇒ Q via WP), WITHOUT showing Q;
%%   (d) ablations: -holdout, -junk, -atom budget, -far probes/jitter
%%       (expect reward inflation + hack families reappearing);
%%   (e) open-book vs closed-book comparison.
Policy-training results are in progress and will be reported in the final version: we train a locally deployable code model with GRPO against the containerized reward service, and evaluate (a) summary tightness on held-out programs, (b) the Frama-C-certified summary rate, and (c) the \emph{downstream} rate at which the closed-book summary alone discharges the benchmark's hidden assertion---the direct test of the claim that summaries subsume targets.
Ablations remove each anti-hacking layer in turn; by construction of the gates in $\S$\ref{sec:exp:gates} we can measure not only reward inflation but the reappearance of specific exploit families.
