%% ============================================================
\section{Introduction}\label{sec:intro}
%% ============================================================

\input{sections/fig_overview}

Loop invariants are the linchpin of deductive program verification.
Ever since the inductive-assertion method of Floyd~\citep{floyd1967assigning} and the axiomatic semantics of Hoare~\citep{hoare1969axiomatic}, verifying a loop $\mathtt{while}\,(B)\,S$ against a contract $\{P\}\cdot\{Q\}$ has reduced to exhibiting an \emph{inductive invariant} $I$ satisfying initiation ($P \Rightarrow I$), consecution ($\{I \wedge B\}\,S\,\{I\}$), and safety ($I \wedge \neg B \Rightarrow Q$).
Finding $I$ automatically has consequently been one of the most intensively studied problems in formal methods.
Four decades of research have attacked it with abstract interpretation~\citep{cousot1977abstract,cousot1978automatic,karr1976affine}, constraint- and template-based synthesis~\citep{colon2003linear,sankaranarayanan2004non,gulwani2008program,gupta2009invgen}, interpolation and IC3/PDR-style generalization~\citep{mcmillan2003interpolation,bradley2011sat,komuravelli2014smt}, abductive inference~\citep{dillig2013inductive}, syntax-guided and stochastic search~\citep{alur2013sygus,fedyukovich2018syntax}, data-driven learning from program states~\citep{sharma2012interpolants,sharma2013data,garg2014ice,garg2016learning,padhi2016pie,zhu2018data,ezudheen2018horn}, neural architectures~\citep{si2018learning,si2020code2inv,ryan2020cln2inv,yao2020gcln}, and, most recently, large language models~\citep{pei2023can,kamath2023finding,chakraborty2023ranking,wu2024lemur,wu2024lam4inv,liu2024acinv,wen2024enchanting,liu2024towards}.

\paragraph{The target-dependence problem.}
Despite their diversity, these lines of work share a common task definition: the input is a program \emph{together with a verification target}---a postcondition, assertion, or safety property $Q$---and the output is an invariant just strong enough to discharge $Q$.
The target is not incidental; it is load-bearing.
CEGIS- and ICE-style learners derive their counterexamples from failed attempts to prove $Q$~\citep{garg2014ice,garg2016learning,padhi2016pie}; Code2Inv shapes its reinforcement signal from the solver's rejection of $Q$-directed verification conditions~\citep{si2018learning,si2020code2inv}; LLM pipelines put the assertion in the prompt and score candidates by whether the verifier closes the goal~\citep{kamath2023finding,chakraborty2023ranking,wu2024lemur,wu2024lam4inv,liu2024acinv}.

We contend that this framing rests on an uncomfortable circularity.
First, verification targets are \emph{scarce}: real code overwhelmingly ships without formal postconditions, and writing them is itself the specification bottleneck that specification mining and, lately, LLM-based specification generation have tried to relieve~\citep{daikon,endres2024can,ma2024specgen,wen2024enchanting}.
Second, and more fundamentally, for loop programs the verification target is very often nothing more than \emph{the invariant itself evaluated at loop exit}.
The safety condition $I \wedge \neg B \Rightarrow Q$ means that the natural $Q$ for a loop---the one a human or a benchmark author writes down---is the exit-time projection of an invariant the author already had in mind; Furia and Meyer make the reverse direction explicit by \emph{deriving} candidate invariants syntactically from postconditions~\citep{furia2010inferring}.
A system that requires $Q$ in order to find $I$ therefore presupposes the very reasoning it is supposed to automate.
Target-directed training compounds the problem: when the reward is ``did the verifier close $Q$,'' the learner is incentivized to produce the \emph{weakest} invariant sufficient for the given target---frequently a light rewriting of $Q$ itself---which transfers to no other property of the same loop.

\paragraph{Our reframing: inductive loop summaries.}
We instead study the task of generating an inductive \emph{loop summary}: given only the precondition, guard, and body---\emph{no postcondition}---produce the strongest practically expressible inductive invariant characterizing the loop's reachable states at the loop head (Figure~\ref{fig:overview}a).
This framing follows the tradition of loop summarization in static analysis and symbolic execution~\citep{kroening2008loop,godefroid2011automatic,xie2016proteus,farzan2015compositional,kincaid2017compositional}, but asks for summaries in the standard inductive-invariant form consumed by deductive verifiers (ACSL, checked by Frama-C/WP~\citep{framac,acsl}).
A tight summary is strictly more useful than a target-directed invariant: it discharges \emph{every} downstream assertion $Q$ with $I \wedge \neg B \Rightarrow Q$, it is reusable across changing specifications, and generating it exercises exactly the semantic reasoning---conservation laws, relational couplings, progress bounds---that the target-directed setting lets a system shortcut.

\paragraph{The reward problem, and why it is hard.}
Dropping the target, however, destroys the training signal every prior learning-based system relied on: with no $Q$ there is no binary verified/failed outcome to optimize.
``Is $I$ inductive?'' alone is useless as a reward---$\mathit{true}$, the guard, and any tautology are inductive.
Recent work concurs that the verifier bit is too coarse a signal for learning even in the target-directed setting: InvBench scores invariants by wall-clock speedup rather than by verification alone~\citep{wei2025invbench}, and \citet{pinto2026notall} show that grading and curating invariants by quality, not mere validity, is what makes them useful training data.
What must be rewarded is \emph{soundness plus empirical tightness}: how effectively does a proved invariant separate sampled reachable states from audited synthetic countertraces?
We operationalize this with an execution-grounded dense reward (Figure~\ref{fig:overview}b,c).
A sampler compiles and runs the instrumented loop over a deterministic input sweep to collect \emph{positive} loop-head states, then constructs three families of synthetic negative candidates: relational perturbations around densely observed trace neighborhoods, continuations of the real body past a genuine exit, and ladder-step escapes beyond a variable's sampled range.
A candidate rollout is first reduced to its inductive core by Houdini-style pruning~\citep{flanagan2001houdini}, with Frama-C/WP as the authoritative backend, and the surviving subset is scored by the fraction of negative traces it rejects, plus a marginal (ablation) term for group-relative credit assignment.

Like any finite-sample signal, such a reward is vulnerable to mislabeled examples and sample-specific predicates~\citep{amodei2016concrete,skalse2022defining,gao2023scaling}.
We therefore keep the trusted mechanism small.  Construction-time vetoes discard a candidate negative if it has been observed reachable or could be a fresh loop entry under admissible inputs, and oracle-controlled loops emit no synthetic negatives; real Houdini then removes candidate invariants that merely fit the finite positive sample but are not inductive.  The scoring rule itself has only two terms: $0.5$ times standalone rejection and $0.5$ times group-ablation marginal rejection.

\paragraph{Contributions.}
This paper makes the following contributions:
\begin{itemize}[leftmargin=1.5em, itemsep=2pt, topsep=2pt]
  \item \textbf{Task reframing.} We identify the target-dependence circularity in loop invariant generation and formulate the alternative task of \emph{closed-book inductive loop summary generation}, which subsumes target-directed generation ($\S$\ref{sec:overview}).
  \item \textbf{An execution-grounded, minimal reward.} We design a dense reward from three audited negative-candidate families and conservative construction guards, with Houdini-pruned survival, trace-unit accounting, and the complete formula $r=0.5\,\mathrm{base}+0.5\,\mathrm{marginal}$ ($\S$\ref{sec:method}).
  \item \textbf{An open three-component pipeline.} \sam{} decomposes into an example sampler, a self-contained HTTP reward service whose image bundles Frama-C/WP, and an independent inference framework that performs no positive/negative sampling and optionally runs verdict-driven multi-round refinement before final certification ($\S$\ref{sec:overview}).
  \item \textbf{Direct validation of discrimination and labels.} The maintained discrimination harness ranks known-quality rollout families on 10 deterministic programs, and the full mislabel audit found zero hard collisions among 486,302 negative witnesses from all 366 benchmark programs ($\S$\ref{sec:experiments}).
\end{itemize}
