%% ============================================================
\section{Method}\label{sec:method}
%% ============================================================

%% ---- Notation ----
\paragraph{Notation.}
For a program $P = (\phi_{\mathrm{pre}}, B, S)$, a \emph{rollout} $A = \{e_1, \dots, e_k\}$ is one candidate clause set produced by one policy sample; a \emph{group} $\mathcal{A} = \{A_1, \dots, A_G\}$ is the set of rollouts sampled for $P$ in one RL step.
The sampler emits an example set $E = (\mathcal{P}, \mathcal{N})$: positives $\mathcal{P}$ (reachable loop-head states) and negatives $\mathcal{N}$ grouped into \emph{trace units}---each unit is one impossible history, witnessed by one or more states.
$\mathrm{surv}(A) \subseteq A$ denotes the subset of $A$ surviving the filter cascade (complexity gate, scope check, positive-soundness, Houdini induction; $\S$\ref{sec:method:reward}).
A clause set \emph{rejects} a trace unit iff some surviving clause evaluates to false on some witness state of that unit; $\mathrm{rej}(X) \subseteq \mathcal{N}$ is the set of units rejected by clause set $X$.

%% ============================================================
\subsection{Sampler: Positives and Truthful Negatives}
\label{sec:method:sampler}

The sampler is the semantic ground truth of the system, and it is deliberately \emph{blind to the postcondition}: it parses only the precondition, guard, and body.

\paragraph{Positives by concrete execution.}
The loop is instrumented with prints at loop entry, at the top of every iteration, and after exit, then compiled with \texttt{gcc} and executed.
Inputs are swept deterministically over a 20-tier magnitude schedule (edge values $0, \pm1, 2, 3$ through moderate magnitudes, clamped to declared per-parameter bounds), in three phases: seed-hashed \emph{far probes} (magnitudes up to 512, cube-safe against 32-bit overflow), \emph{stripes} (all parameters move together), and a mixed-radix \emph{grid} (tier combinations, so parameter-coupling preconditions such as $z = k$ receive satisfying tuples).
Every tuple is checked against the full precondition before use---out-of-contract inputs would poison the positive set.
Nondeterministic calls (\texttt{unknown()}) are auto-defined as seeded random oracles so repeated runs explore distinct branches.
Loops run to their \emph{real} exit; print throttling keeps long loops cheap while keeping trace windows dense, and a run that hits the iteration cap is flagged so that no range conclusion is drawn from a truncated trace.

\paragraph{Truthful negatives.}
A negative that is actually reachable would penalize exactly the true invariants, so every negative must be \emph{provably impossible}---truthful by construction, derived from the loop's own dynamics rather than from executing mutated code.
Negatives come in two balanced dimensions:

\begin{itemize}[leftmargin=1.5em, itemsep=2pt, topsep=2pt]
\item \textbf{Relational negatives} expose missing laws ($x + y = n$, $z = x \cdot y$).
Small single-axis and pairwise perturbations ($\pm\{1,2,3\}$) are applied to \emph{dense bases}: trace states whose next $w = 4$ successors were all sampled, so every reachable in-window neighbor is known and a surviving perturbation is genuinely off the reachable manifold.
\item \textbf{Bound negatives} expose missing range facts.
\emph{Over-run states} execute the real body up to 24 steps past a genuine exit: they preserve every (even non-linear) relational law while violating the range, and each over-run continuation forms \emph{one} trace unit.
\emph{Box-escape states} take large ladder steps ($\pm\{5, 8, 13, 21, 34, 55, 89\}$) kept only when they leave the variable's observed range.
\end{itemize}

Perturbed states that could constitute a fresh loop entry (parameters at their pre-values, literal-initialized locals at their initializers, precondition satisfied) are reachable and are dropped.
Under nondeterminism the sampler degrades soundly rather than optimistically: a nondeterministic guard disables the entire bound dimension (the loop can always run longer); nondet-tainted variables are never perturbed directly, but sound signal is \emph{rescued} for them via conserved rigid pairs ($c_w v - c_v w$), gcd-lattice residues ($v \equiv v_0 \pmod g$), monotone floors/ceilings, and guard-implied bounds---each a provably sound fact even when no reachable range exists.

\paragraph{Trace-unit accounting.}
Negatives are counted per impossible \emph{history}, not per state: an over-run trajectory contributes one unit no matter how many witness states it carries, and a unit is rejected if any witness is.
This prevents count inflation from long continuations and makes the reward's denominator semantically meaningful.

%% ============================================================
\subsection{Reward: Dense Scoring of Summary Tightness}
\label{sec:method:reward}

Given a program, a group $\mathcal{A}$, and an example set $E$, each rollout is scored in three steps.

\paragraph{Step 1: survival.}
$A \mapsto \mathrm{surv}(A)$ applies, in order: a per-clause \emph{complexity gate} (clauses longer than 160 characters or with more than 12 comparison atoms are dropped, not truncated); a scope check (clauses referencing out-of-scope identifiers are dropped); \emph{positive soundness} (any clause false on some sampled positive is dropped, with disequalities $v \ne c$ additionally checked against the \emph{full} observed value set of $v$ rather than a subsample); and, when Frama-C is available, Houdini pruning to the greatest inductive subset~\citep{flanagan2001houdini}.
Positive subsampling always retains per-variable extremes---the witnesses that falsify overfitted boxes.
Survival results are memoized on the normalized clause set, since ablation re-filters heavily overlapping sets.

\paragraph{Step 2: per-rollout scores.}
With $n = |\mathcal{N}|$ trace units and $U = \bigcup_{A \in \mathcal{A}} A$ the group union,
\begin{align}
\mathrm{base}(A) &= \frac{|\mathrm{rej}(\mathrm{surv}(A))|}{n},
&
\mathrm{marg}(A) &= \frac{\bigl|\mathrm{rej}(\mathrm{surv}(U)) \setminus \mathrm{rej}(\mathrm{surv}(U \setminus A))\bigr|}{n},
\end{align}
i.e.\ base measures how much of the impossible space the rollout alone excludes, and the marginal term is the rollout's ablation gain over the rest of the group---group-relative credit in the spirit of GRPO~\citep{shao2024deepseekmath}, computed against the verifier-pruned union rather than against raw scores.
A clause is \emph{useful} if it uniquely rejects at least one unit no other clause of the rollout rejects; with $\mathrm{junk}(A) = 1 - \mathrm{useful}(A)/|A|$, the reward is
\begin{equation}
r(A) \;=\; \max\bigl(0,\; w_b \,\mathrm{base}(A) + w_m \,\mathrm{marg}(A) - w_j \,\mathrm{junk}(A)\bigr),
\qquad (w_b, w_m, w_j) = (0.5,\, 0.5,\, 0.05).
\label{eq:reward}
\end{equation}
The group additionally receives $\mathrm{batch} = |\mathrm{rej}(\mathrm{surv}(U))| / n$ and a re-roll flag when $\mathrm{batch} < 0.6$.

\paragraph{Step 3: multi-seed minimum.}
Scores are computed against $s{=}2$ independent example sets plus $h{=}1$ \emph{holdout} set drawn with a fresh seed per reward call, and each rollout's final $\mathrm{base}/\mathrm{marg}/r$ is the \emph{minimum} across sets.
A true invariant is indifferent to the sampling seed; anything fitted to one concrete sample is not, and the minimum ensures memorizing any single deterministic sample earns nothing.

Negative rejection is evaluated by a locked-down predicate evaluator with C integer semantics (truncation-toward-zero division and remainder, matching the compiled traces), so the reward needs no SMT call on its hot path; the only heavyweight cost is Houdini, bounded by $G{+}1$ pruner runs per group and cut further by memoization and a lite-to-full filter cascade.

%% ============================================================
\subsection{Anti-Hacking Defenses}
\label{sec:method:antihack}

Because the reward is computed from a finite sample, ``memorize the sample'' is always an available policy; we enumerate the attack surface and close each vector, treating hack resistance as a specification with its own regression gates ($\S$\ref{sec:experiments}).

\begin{itemize}[leftmargin=1.5em, itemsep=3pt, topsep=2pt]
\item \textbf{Mega-conjunction} (one gate-compliant clause enumerating the negative pool): blocked by the per-clause complexity gate (160 chars / 12 atoms, drop-not-truncate).
\item \textbf{Split spray} (hundreds of small $v \ne c$ clauses): blocked by a \emph{rollout atom budget}---only the emission-order prefix within 32 total atoms is scored.
Honest summaries (3--6 clauses of 1--4 atoms) fit with headroom; a spray sacrifices its budget for clauses the next defense zeroes out.
\item \textbf{Gold-plus-spray} (ride one true clause, pad with shadows): the junk term of Eq.~\ref{eq:reward} discounts clauses that reject nothing uniquely, so padding strictly lowers reward.
\item \textbf{Sample-extreme boxes} ($lo \le v \le hi$ read off the observed range): the holdout seed's \emph{far probes} give each seed a different input envelope, so a box fitted to canonical extremes is falsified by holdout positives (dropped in survival) or misses holdout negatives; a symbolic bound ($x \le n$) is envelope-indifferent.
\item \textbf{Pointwise memorization} ($v \ne c$ farms aimed at negative clusters): \emph{tier jitter} (seed-hashed $\pm2$ on non-edge tiers, uniform within a run so coupled preconditions stay satisfiable) decorrelates sampled values across seeds, and bound-negative offsets are themselves seed-hashed; the cross-seed minimum then zeroes the farm.
\item \textbf{Slice concentration} (many negatives sharing one farm-targetable coordinate): a slice cap ($\le 6$ units per never-seen-positive coordinate value) bounds what any 32-atom farm can reach.
\item \textbf{Dimension starvation} (score $\approx 1$ on relations while ignoring bounds): relation units are capped at twice the bound units, so bounds carry a program-independent share of the reward.
\end{itemize}

The defenses compose: capacity limits make memorization expensive, uniqueness accounting makes it unpaid, and cross-seed minimum scoring makes it worthless even if afforded.
$\S$\ref{sec:experiments} validates the composition adversarially.

%% ============================================================
\subsection{Inference: Closed-Book Generation with Certified Repair}
\label{sec:method:inference}

At inference time the policy is prompted with the program \emph{after postcondition stripping} (closed-book): the prompt asks for the strongest inductive ACSL loop invariants capturing the loop's behavior---conservation and relational laws plus tight progress bounds---emitted only as \texttt{loop invariant} lines.
The system prompt encodes the Frama-C/WP practicalities: one core relational equality, minimal bounds, exact \texttt{loop assigns}, no tautologies, non-strict weakening of strict guards at exit, and pre-state references via \texttt{\textbackslash at(v,Pre)} for parameters only.

For each program the framework samples $n$ rollouts (cloud API or in-process batched vLLM), unions and deduplicates their clauses, and applies the same survival cascade as the reward---ending, here, in full Houdini pruning and Frama-C/WP verification of the re-annotated program.
Repair is thus Houdini-style subtraction rather than conversational self-correction: unsound clauses are dropped by the pruner, not re-prompted.
If verification fails, the framework re-rolls up to a bounded number of attempts, keeping the best result by (verified, summary size).
Training and inference share one canonical sampler configuration, so the reward optimized during RL is exactly the quantity the deployed pipeline measures.
