%% ============================================================
\section{Problem Formulation and System Overview}\label{sec:overview}
%% ============================================================

\subsection{From Target-Directed Invariants to Loop Summaries}
\label{sec:overview:task}

Fix a loop program $P = (\phi_{\mathrm{pre}}, B, S)$ over integer variables $V$: a precondition $\phi_{\mathrm{pre}}$, a guard $B$, and a body $S$.
A \emph{state} $\sigma : V \to \mathbb{Z}$ is a valuation of $V$ (together with the pre-state values $v_0$ of parameters, so that summaries may relate current to initial values).
Let $\mathcal{R}(P)$ denote the set of states reachable \emph{at the loop head} from some initial state satisfying $\phi_{\mathrm{pre}}$.

\begin{definition}[Inductive invariant]
A predicate $I$ over $V$ is an inductive invariant of $P$ if
(i) \emph{initiation}: every loop-head state reached directly from $\phi_{\mathrm{pre}}$ satisfies $I$; and
(ii) \emph{consecution}: $\{I \wedge B\}\, S\, \{I\}$.
Equivalently, $I \supseteq \mathcal{R}(P)$ and $I$ is closed under the loop's transition relation.
\end{definition}

\begin{definition}[Target-directed generation]
Given $(P, Q)$, produce any inductive $I$ with $I \wedge \neg B \Rightarrow Q$.
\end{definition}

\begin{definition}[Inductive loop summary generation (ours)]
Given $P$ alone, produce an inductive $I$ that is as \emph{tight} as practically expressible: among predicates in the assertion language, $I$ should exclude as much of the unreachable space $\overline{\mathcal{R}(P)}$ as possible.
The ideal object is the strongest inductive invariant, i.e.\ (the assertion-language best approximation of) $\mathcal{R}(P)$ itself.
\end{definition}

The summary task subsumes the target-directed one: any $Q$ provable from \emph{some} inductive invariant is provable from the strongest one.
It is also strictly harder to game.
In the target-directed task, $Q$ itself (weakened to pass initiation) is often an acceptable answer; in the summary task there is no $Q$ to parrot, and weak answers---$\mathit{true}$, the guard, loose boxes---are exactly what the reward of $\S$\ref{sec:method} is designed to score low.

In practice a summary is a set of ACSL \texttt{loop invariant} clauses whose conjunction is the invariant.  Typical summaries pair one or two relational laws (e.g.\ $x = y \cdot z$ preserved by the body) with progress bounds (e.g.\ $0 \le i \le n$).
Soundness is never entrusted to the learner in production: every clause set is reduced to its greatest inductive subset by Houdini pruning~\citep{flanagan2001houdini} over Frama-C/WP~\citep{framac} before it is scored or reported.  A native installation without Frama-C can fall back to the positive filter for development, but that path is not a proof certificate; both deployment images bundle the prover.

\subsection{System Overview}
\label{sec:overview:system}

\sam{} decomposes into three operationally independent components.  Training calls the reward service, while deployment evaluation runs inference; neither invokes the other:

\begin{stepbox}[1. Example sampler (\S\ref{sec:method:sampler})]
Compiles and executes the instrumented loop over a deterministic, precondition-respecting input sweep; emits sampled positive loop-head states and three families of synthetic negative-candidate trace units: relational perturbations, genuine-exit over-runs, and sampled-range escapes.
The sampler sees only $(\phi_{\mathrm{pre}}, B, S)$---never the postcondition.
\end{stepbox}

\begin{stepbox}[2. Reward framework (\S\ref{sec:method:reward})]
Scores a \emph{group} of rollouts (candidate clause sets sampled for the same program, as in group-relative RL~\citep{shao2024deepseekmath}).
Each rollout is reduced to its surviving inductive core, then receives $0.5$ times the fraction of negative traces it rejects alone (base) plus $0.5$ times its ablation gain in the group union (marginal).  This is the entire reward formula.
Packaged as a self-contained HTTP service (\texttt{POST /reward}) with gcc, Frama-C/WP, Why3, and Z3 bundled in the image, plus an offline JSONL/Parquet scorer.
\end{stepbox}

\begin{stepbox}[3. Inference framework (\S\ref{sec:method:inference})]
Prompts the policy \emph{closed-book} by default, unions and deduplicates candidate rollouts, optionally runs up to $m$ rounds of WP-verdict feedback and LLM refinement, then applies full Houdini and Frama-C/WP verification.
It is independent of the reward and never invokes the execution sampler or constructs positive/negative examples; a vLLM rollout provider supplies model candidates.
\end{stepbox}

Two evaluation harnesses guard the implemented contract: discrimination over 10 deterministic programs with hand-written quality specifications, and a full-suite negative-mislabel audit ($\S$\ref{sec:experiments}).

\paragraph{Running example.}
Consider a benchmark loop (abridged from our linear suite):
\begin{lstlisting}
void foo(int n) {  // requires n >= 0
  int x = 0, y = n;
  while (y > 0) { x = x + 1; y = y - 1; }
}
\end{lstlisting}
The sampler observes loop-head states such as $(x{=}0, y{=}5)$, $(x{=}1, y{=}4)$, \dots for $n{=}5$ and many other inputs.
Relational negatives perturb dense trace neighborhoods---$(x{=}2, y{=}4)$ with $n{=}5$ breaks $x + y = n$; bound negatives continue past the real exit---$(x{=}6, y{=}{-}1)$ preserves $x+y=n$ yet is impossible.
The summary $\{x + y = n,\; 0 \le y \le n\}$ (with $n \ge 0$ from the precondition) rejects every constructed negative trace, whereas the guard restatement $y \ge 0$ rejects only a subset of the bound negatives and a tautology rejects none.
Any exit-time fact a user later cares about---$x = n$, $y = 0$, $x \ge y$---follows from the summary, though the generator never saw such a target.
